\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  urlcolor=blue!60!black,
  citecolor=blue!60!black,
  pdftitle={Previo Technical Report},
  pdfauthor={Engineering Team},
  pdfsubject={Collaborative Markdown and LaTeX Writing Platform}
}

\definecolor{codebg}{RGB}{247,247,247}
\definecolor{codecomment}{RGB}{0,120,0}
\definecolor{codekeyword}{RGB}{0,0,160}
\definecolor{codestring}{RGB}{163,21,21}

\lstdefinestyle{previo}{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{codebg},
  commentstyle=\color{codecomment},
  keywordstyle=\color{codekeyword}\bfseries,
  stringstyle=\color{codestring},
  breaklines=true,
  frame=single,
  columns=fullflexible,
  keepspaces=true
}

\setstretch{1.15}
\titleformat{\section}{\large\bfseries}{\thesection}{0.75em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.75em}{}

\pagestyle{fancy}
\fancyhf{}
\lhead{Previo Technical Report}
\rhead{\today}
\cfoot{Page \thepage\ of \pageref{LastPage}}

\title{
  \vspace{-1cm}
  \textbf{Previo: Technical Report}\\
  \large Collaborative Markdown and LaTeX Writing Platform
}
\author{The Previo Team}
\date{}

\begin{document}
\maketitle
\newpage
\tableofcontents
\newpage

\section{Executive Summary}
Previo is a collaborative writing platform designed to support Markdown and LaTeX workflows in a single web application. The platform combines an interactive editor, live preview rendering, project collaboration, invitation-based access control, document versioning, and per-project workspace file management. The system is implemented with a Next.js frontend/backend gateway and Python rendering/data services, backed by PostgreSQL.

The primary goal is to provide an integrated environment where users can:
\begin{itemize}[leftmargin=1.5em]
  \item create and manage technical writing projects,
  \item collaborate with peers through controlled project sharing,
  \item render and export Markdown/LaTeX artifacts,
  \item maintain reliable revision history with revert capabilities,
  \item store project-scoped assets (images/files) for reproducible documents.
\end{itemize}

\section{Problem Statement and Objectives}
\subsection{Problem Statement}
Technical writing teams often rely on fragmented tooling: separate editors, versioning workflows, collaboration channels, and rendering pipelines. This increases context switching and operational complexity, especially for LaTeX-heavy documents requiring compile toolchains and asset management.

\subsection{Objectives}
The project was designed around the following objectives:
\begin{enumerate}[leftmargin=1.5em]
  \item Provide a unified writing workspace for Markdown and LaTeX.
  \item Support secure authentication and session management.
  \item Enable project-level collaboration with invitation workflows.
  \item Deliver reliable rendering/preview for both formats.
  \item Track full document history and support selective revert.
  \item Keep deployment straightforward via Docker Compose and CLI automation.
\end{enumerate}

\section{System Overview}
\subsection{High-Level Architecture}
The platform follows a multi-layer architecture:
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Presentation Layer (Next.js UI):} editor, dashboard, authentication pages, project management screens.
  \item \textbf{Application API Layer (Next.js API Routes):} request validation, authorization, orchestration, and file handling.
  \item \textbf{Service Layer (Python scripts):} Markdown/LaTeX rendering and PostgreSQL data operations.
  \item \textbf{Persistence Layer (PostgreSQL + filesystem):} relational data and project workspace assets.
\end{itemize}

\subsection{Core Modules}
\begin{longtable}{p{0.28\linewidth} p{0.66\linewidth}}
\toprule
\textbf{Module} & \textbf{Responsibility} \\
\midrule
Auth & Signup/login, OAuth integration, session token lifecycle \\
Projects & Create/read/update/delete projects, search, formatting metadata \\
Collaboration & Invitations, membership management, collaborator visibility \\
Versioning & Save snapshots, generate diffs, restore previous states \\
Rendering & Markdown HTML preview/export, LaTeX PDF preview/export \\
Workspace & Upload/list/open/delete project assets; asset linking in documents \\
CLI/DevOps & Start/stop workflow using cross-platform command wrappers \\
\bottomrule
\end{longtable}

\section{Technology Stack}
\subsection{Frontend and API}
\begin{itemize}[leftmargin=1.5em]
  \item Next.js (App Router)
  \item React + TypeScript
  \item Tailwind CSS
  \item API routes for backend orchestration
\end{itemize}

\subsection{Backend Services}
\begin{itemize}[leftmargin=1.5em]
  \item Python 3 scripts for rendering and data access
  \item PostgreSQL service script handling migrations and business logic
  \item LaTeX toolchain via \texttt{pdflatex}
\end{itemize}

\subsection{Infrastructure}
\begin{itemize}[leftmargin=1.5em]
  \item Docker Compose for local and production-like orchestration
  \item Named volumes for PostgreSQL persistence and workspace storage
  \item Node-based CLI wrapper for lifecycle commands
\end{itemize}

\section{Data Model}
\subsection{Primary Entities}
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{users:} identity, profile, credential hash, creation timestamp
  \item \textbf{sessions:} token, user linkage, expiration, login state
  \item \textbf{projects:} owner, title, format, content, update timestamp
  \item \textbf{project\_members:} accepted collaborators
  \item \textbf{project\_invitations:} pending/accepted invitation records
  \item \textbf{project\_versions:} immutable snapshots and diff metadata
\end{itemize}

\subsection{Consistency and Migration Strategy}
Database schema setup and backward-compatible evolution are performed during service startup. The migration logic checks for missing columns/tables and applies idempotent updates to avoid destructive resets.

\section{API Design}
\subsection{Authentication APIs}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{/api/auth/signup}
  \item \texttt{/api/auth/login}
  \item \texttt{/api/auth/logout}
  \item \texttt{/api/auth/oauth/\{provider\}}
\end{itemize}

\subsection{Project and Collaboration APIs}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{/api/projects} (list/create)
  \item \texttt{/api/projects/\{id\}} (read/update/delete)
  \item \texttt{/api/projects/import} (create project from \texttt{.md}/\texttt{.tex})
  \item \texttt{/api/projects/\{id\}/collaborators}
  \item \texttt{/api/invitations} (list/respond)
\end{itemize}

\subsection{Versioning and Workspace APIs}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{/api/projects/\{id\}/versions}
  \item \texttt{/api/projects/\{id\}/versions/revert}
  \item \texttt{/api/projects/\{id\}/workspace}
  \item \texttt{/api/projects/\{id\}/workspace/file}
\end{itemize}

\section{Implementation Details}
\subsection{Editor and Preview Workflow}
The editor supports two modes: Markdown and LaTeX. Content changes trigger debounced preview rendering requests. Markdown previews return HTML payloads; LaTeX previews return PDF data URLs for iframe embedding.

\subsection{Versioning Logic}
Each save operation creates a new immutable snapshot capturing:
\begin{itemize}[leftmargin=1.5em]
  \item title, format, and content at save time,
  \item actor identity,
  \item change summary (explicit or generated),
  \item unified diff for audit and inspection.
\end{itemize}

\subsection{Workspace Asset Handling}
Each project has a dedicated workspace directory. The workspace module enforces safe path resolution and blocks path traversal attacks. Uploaded files are normalized and collision-safe naming is applied.

For LaTeX rendering, workspace image URLs included in source are resolved to local temporary assets before invoking \texttt{pdflatex}. This guarantees compatibility with LaTeX compilers, which operate on local files rather than HTTP query paths.

\subsection{LaTeX Resilience Enhancements}
The rendering pipeline sanitizes common placeholder patterns found in copied templates (e.g., invalid figure placement and graphics option placeholders) before compilation. This reduces avoidable user-facing compile failures.

\section{Security Considerations}
\subsection{Authentication and Session Security}
\begin{itemize}[leftmargin=1.5em]
  \item Passwords are hashed before storage.
  \item Session tokens are random, expiration-bound, and server-validated.
  \item Unauthorized API calls return explicit 401 responses.
\end{itemize}

\subsection{Authorization}
\begin{itemize}[leftmargin=1.5em]
  \item Project access checks are enforced on read/write endpoints.
  \item Deletion is restricted to project owners.
  \item Invitation acceptance is constrained by recipient email identity.
\end{itemize}

\subsection{File Safety}
\begin{itemize}[leftmargin=1.5em]
  \item Workspace paths are canonicalized and validated.
  \item File access remains scoped to project workspace roots.
  \item API routes enforce auth before file operations.
\end{itemize}

\section{Deployment and Operations}
\subsection{Containerization}
The platform is deployed as Docker Compose services:
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{postgres} for relational persistence
  \item \textbf{ui} for Next.js runtime + Python scripts + LaTeX toolchain
\end{itemize}

\subsection{Runtime Volumes}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{postgres\_data}: durable database storage
  \item \texttt{workspaces\_data}: project asset storage
\end{itemize}

\subsection{CLI-Oriented Workflow}
The \texttt{previo} command standardizes lifecycle operations:
\begin{lstlisting}[style=previo,language=bash]
previo start 3000
previo stop
\end{lstlisting}
This reduces setup friction across Linux, macOS, and Windows environments.

\section{Testing and Validation}
\subsection{Functional Validation Areas}
\begin{itemize}[leftmargin=1.5em]
  \item Auth flows: signup/login/logout/OAuth
  \item Project CRUD and search
  \item Collaboration invitation lifecycle
  \item Version snapshot creation and revert correctness
  \item Markdown and LaTeX preview/export stability
  \item Workspace file upload/open/delete behavior
\end{itemize}

\subsection{Error Handling Validation}
The system validates and surfaces:
\begin{itemize}[leftmargin=1.5em]
  \item invalid payloads and missing fields (400),
  \item unauthorized access (401),
  \item not-found resources (404),
  \item backend/render failures with diagnostic context (500).
\end{itemize}

\section{Performance and Scalability Notes}
\subsection{Current Performance Characteristics}
\begin{itemize}[leftmargin=1.5em]
  \item API layer is lightweight and script-driven.
  \item Rendering cost is dominated by LaTeX compilation and document size.
  \item Project listing and search rely on SQL filtering and sort by update time.
\end{itemize}

\subsection{Scalability Opportunities}
\begin{itemize}[leftmargin=1.5em]
  \item move heavy rendering to background jobs/worker queues,
  \item add indexed full-text search for large project datasets,
  \item adopt object storage for workspace assets at scale,
  \item add real-time collaboration transport (e.g., WebSocket/CRDT layer).
\end{itemize}

\section{Risks and Mitigations}
\begin{longtable}{p{0.28\linewidth} p{0.30\linewidth} p{0.36\linewidth}}
\toprule
\textbf{Risk} & \textbf{Impact} & \textbf{Mitigation} \\
\midrule
LaTeX compile instability & Preview/export failures & Source sanitization, clearer errors, controlled temp workspace \\
Large file uploads & Storage and latency pressure & Upload limits, per-project scoping, cleanup policies \\
Unauthorized access attempts & Data leakage risk & Strict auth checks and project membership validation \\
Schema drift across environments & Runtime faults & Startup migration checks and idempotent DDL \\
\bottomrule
\end{longtable}

\section{Roadmap}
\subsection*{Near-Term}
\begin{itemize}[leftmargin=1.5em]
  \item richer project templates and onboarding
  \item improved inline diagnostics for LaTeX compile errors
  \item stronger automated test coverage (API + integration)
\end{itemize}

\subsection*{Mid-Term}
\begin{itemize}[leftmargin=1.5em]
  \item real-time collaborative editing primitives
  \item role-based project permissions beyond owner/member
  \item asynchronous rendering jobs with progress feedback
\end{itemize}

\subsection*{Long-Term}
\begin{itemize}[leftmargin=1.5em]
  \item enterprise SSO options
  \item advanced observability and SLO dashboards
  \item multi-tenant hardening and policy controls
\end{itemize}

\section{Conclusion}
Previo provides a practical and extensible foundation for collaborative technical writing across Markdown and LaTeX. By unifying editing, rendering, versioning, collaboration, and workspace management, it reduces operational friction for teams producing technical documents. The architecture balances rapid delivery with clear upgrade paths for scale, reliability, and enterprise readiness.

\section*{Appendix A: Sample Environment Variables}
\addcontentsline{toc}{section}{Appendix A: Sample Environment Variables}
\begin{lstlisting}[style=previo]
POSTGRES_DB=previo
POSTGRES_USER=previo
POSTGRES_PASSWORD=previo
PREVIO_DATABASE_URL=postgresql://previo:previo@postgres:5432/previo
NEXT_PUBLIC_APP_URL=http://localhost:3000
APP_URL=http://localhost:3000
PREVIO_WORKSPACES_DIR=/app/workspaces
GOOGLE_CLIENT_ID=
GOOGLE_CLIENT_SECRET=
GITHUB_CLIENT_ID=
GITHUB_CLIENT_SECRET=
\end{lstlisting}

\section*{Appendix B: Build and Run Commands}
\addcontentsline{toc}{section}{Appendix B: Build and Run Commands}
\begin{lstlisting}[style=previo,language=bash]
# Install CLI
npm link

# Start stack on port 3000
previo start 3000

# Stop stack
previo stop
\end{lstlisting}

\end{document}
